---
title: "Statistical Inference hw2"
author: "Max Gribov"
date: "Sunday, January 25, 2015"
output: html_document
---
### Analyze the ToothGrowth data in the R datasets package.  

#### Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
# load the data
library(ggplot2)
library(datasets)
data <- ToothGrowth
```

##### Data Exploration  
**How does the dose affect total tooth length**
```{r}
ggplot(
  data=data, aes(x=as.factor(dose), y=len, fill=supp)) +
  geom_bar(stat="identity",) +
  facet_grid(. ~ supp) +
  xlab("Dose in miligrams") +
  ylab("Tooth length") +
  guides(fill=guide_legend(title="Supplement type")
)
```
  
  
**How does supplement type affect the tooth length**
```{r}
ggplot(
  data=data, aes(x=as.factor(supp), y=len, fill=dose)) +
  geom_bar(stat="identity",) +
  facet_grid(. ~ dose) +
  xlab("Supplement Type") +
  ylab("Tooth length") +
  guides(fill=guide_legend(title="Dose in Miligrams")
)

```
   
##### Conclusion 1: There appears to be a stronger correlation between the Dose in Miligrams and the Total Length of the tooth.


#### Provide a basic summary of the data.
```{r}
summary(data)
```


#### Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.   
**Use t-test to see if the Supplement Type has significant effect on Tooth Length**
```{r}
t.test(len ~ supp, data = data)
```
##### Conclusion 2:  The supplement type does not seem to have signigicant effect on Tooth Length. p-value is larger than 5%, and the Confidence Interval includes 0.  
  
**Use t-test to see if the Dose In Miligrams has significant effect on Tooth Length**
```{r}
# we have 3 dosage levels
dose_5_1 <- rbind(subset(data, dose == 0.5), subset(data, dose == 1.0))
dose_5_2 <- rbind(subset(data, dose == 0.5), subset(data, dose == 2.0))
dose_1_2 <- rbind(subset(data, dose == 1.0), subset(data, dose == 2.0))

# each test will examine smaller vs bigger dosages
t.test(len ~ dose, data = dose_5_1)
t.test(len ~ dose, data = dose_5_2)
t.test(len ~ dose, data = dose_1_2)
```  
##### Conclusion 3:  The supplement does does seem to have ad effect on Tooth Length. All of p-values are small, and the Confidence Intervals do not include 0.  
  

#### State your conclusions and the assumptions needed for your conclusions.  
It appears that the tooth length is affected by the size of dose in miligrams much more than by the supplement type.    
  
Assumptions:  
Members of the sample population represent the entire population.  
Members of the sample population were randomly assigned the dosages and supplement types.  
